We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2017-12-03 paper/paper.tex
covariance data illustris
illustrisdm cov
elvis cov
updated abstract

\begin{abstract}
Here we quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We build explicit probability distributions of finding an aspherical
satellite distribution. 
This allows us to estimate the atipicallity of the satellite
distributions in the LG even when the underlying simulations
do not have enough systems that resemble the LG features.
This is also standarized method to compare different simulations and
is also convenient for reproducibility and comparison 
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-DarkMatter and ELVIS simulations.
Although there are significant differences between the three
simulations, we find that at most of $0.20\%$ the pairs are expected
to have satellite distributions with the same degree of deviation from
sphericity as the Local Group.
The three simulations show that the satellite
distribution in M31 is common while the very planar MW satellite
distribution is hard to find; we expect that between $19\%$ to $24\%$
of the pairs have halo similar to M31, while only $0.80\%$ are similar
to the MW.

\end{abstract}

@@

\section{Covariance Matrices and Mean value vectors}
Illustris
\begin{verbatim}
[[ 0.8344  0.7849  0.4186 -0.1155  0.0340 -0.1614]
 [ 0.7849  1.1698 -0.1568 -0.0793  0.0362 -0.1213]
 [ 0.4186 -0.1568  0.9606 -0.1151 -0.0150 -0.1098]
 [-0.1155 -0.0793 -0.1151  0.7873  0.6339  0.4006]
 [ 0.0340  0.0362 -0.0150  0.6339  0.6932  0.0652]
 [-0.1614 -0.1213 -0.1098  0.4006  0.0652  0.6100]]
\end{verbatim}

\begin{verbatim}
[-0.1852 -0.4152    -0.2002 -0.1786 -0.4072 -0.2394]
\end{verbatim}

ELVIS
\begin{verbatim}
[[ 1.6051  1.2207  0.7062  0.0833  0.1063  0.0812]
 [ 1.2207  1.0872  0.3584  0.0298  0.1085 -0.0324]
 [ 0.7062  0.3584  0.6363  0.0904  0.0194  0.1601]
 [ 0.0833  0.0298  0.0904  0.4534  0.6138 -0.1159]
 [ 0.1063  0.1085  0.0194  0.6138  1.2120 -0.6372]
 [ 0.0812 -0.0324  0.1601 -0.1159 -0.6372  0.6475]]
\end{verbatim}

\begin{verbatim}
[-0.1803 -0.2829 -0.2205 -0.3292 -0.7484 -0.0863]
\end{verbatim}
\end{document}

>>>>

\begin{abstract}
We quantify the asphericity in the spatial distribution of the
brightest satellites around the Milky Way (MW) and M31, the dominant
galaxies in the Local Group (LG).
We use this approach on simulations to build an explicit probability
distribution of finding an aspherical  satellite  distribution around
galaxies similar to the Milky Way and M31. 
This probability distribution allows us to estimate the atipicallity
of the satellite distributions in the LG even when the underlying
simulations do not have enough systems that resemble the LG satellite
anisotropy. 
This methodology is straightforward to implement, allowing
for an straighforward quantification of aspherical satellite
distributions both in simulations and observations.
We demonstrate the method using three different simulations:
Illustris-1,  Illustris-1-DarkMatter and ELVIS simulations. 
We find as a common trend that at most of $0.20\%$ the pairs are
expected  to have satellite distributions with the same degree of
deviation from sphericity as the Local Group.
The three simulations show that the satellite distribution in M31 is
common while the very planar MW satellite distribution is hard to
find.
We show that in these simulations between $19\%$ to $24\%$ of the pairs
have a halo with a satellite distribution similar to M31, while only
$0.80\%$, at most, have a satellite distribution as planar as the MW.
This characteristic places the LG as at the level of a 3-$\sigma$
outlier in context provided by LCDM simulations.
\end{abstract}

@@

\section{Covariance Matrices and Mean value vectors}
Illustris M31

Covariance
\begin{verbatim}
[[ 0.83+/-0.03  0.78+/-0.04  0.41+/-0.04]
 [ 0.78+/-0.04  1.16+/-0.05 -0.15+/-0.05]
 [ 0.41+/-0.04 -0.15+/-0.05  0.96+/-0.04]]
\end{verbatim}

Mean

\begin{verbatim}
[-0.18+/-0.04 -0.41+/-0.05  -0.20+/-0.05]
\end{verbatim}

Illustris MW
\begin{verbatim}
[[ 0.78+/-0.06  0.63+/-0.07  0.40+/-0.03]
 [ 0.63+/-0.07  0.69+/-0.08   0.06+/-0.02]
 [ 0.40+/-0.03  0.06+/-0.02   0.61+/-0.04]]
\end{verbatim}

Mean
\begin{verbatim}
[-0.17+/-0.04 -0.40+/-0.04 -0.23+/-0.04]
\end{verbatim}

IllustrisDM M31

Covariance
\begin{verbatim}
[[ 1.50+/-0.08  1.27+/-0.08  0.64+/-0.04]
 [ 1.27+/-0.08  1.31+/-0.08  0.28+/-0.04]
 [ 0.64+/-0.04  0.28+/-0.04  0.70+/-0.04]]
\end{verbatim}

Mean
\begin{verbatim}
[-0.37+/-0.05 -0.55+/-0.05  -0.13+/-0.03]
\end{verbatim}

IllustrisDM M31

Covariance
\begin{verbatim}
[[ 1.03+/-0.05  0.94+/-0.05  0.36+/-0.03 ]
 [ 0.94+/-0.05  1.24+/-0.07 -0.18+/-0.04]
 [ 0.36+/-0.03  -0.18+/-0.04  0.86+/0.03]]
\end{verbatim}

Mean
\begin{verbatim}
[-0.45+/-0.04 -0.61+/-0.05 -0.31+/-0.04]
\end{verbatim}

ELVIS M31

Covariance
\begin{verbatim}
[[ 1.60+/-0.19  1.22+/-0.14  0.70+/-0.09]
 [ 1.22+/-0.14  1.08+/-0.10  0.35+/-0.05]
 [ 0.70+/-0.09  0.35+/-0.05  0.63+/-0.10]]
\end{verbatim}

Mean
\begin{verbatim}
[-0.18+/-0.08 -0.28+/-0.07 -0.22+/-0.05]
\end{verbatim}

ELVIS MW
Covariance
\begin{verbatim}
[[ 0.45+/-0.04    0.61+/-0.09 -0.11+/-0.09 ]
 [ 0.61+/-0.09  1.21+/-0.17 -0.63+/-0.09]
 [-0.11+/-0.05  -0.63+/-0.09  0.64+/-0.06]]
\end{verbatim}

\begin{verbatim}
[-0.32+/-0.04 -0.74+/-0.07 -0.08+/-0.05]
\end{verbatim}

\end{document}
